---
title: "Final Dataproject"
author: "skyggemenn inc^TM."
date: "24/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls())

library(dplyr)
library(ggplot2)
library(tidyverse)
library(gapminder)
library(gganimate)
library(hrbrthemes) # ikke sikker på om denne trenger å være med 
library(png)
library(gifski)
library(magick)
library(broom)
library(plotly)
library(crosstalk)
library(R2HTML)
packageVersion("plotly")


```

```{r, include=FALSE}
#Laster opp og rydder HDI filen
hdi <- read.csv("human-development-index.csv")
hdi <- rename(hdi, "Country" = "Entity")
hdi <- rename(hdi, "Value" = "X..0.1..higher.values.are.better..")
hdi <- hdi[ ,-2]
```

```{r, include=FALSE}
#Laster opp og rydder renewable energy filen

energy <- read.csv("modern-renewable-energy-consumption.csv")
energy <- rename(energy, "Country" = "Entity")
energy <- rename(energy, "Hydro" = "Hydropower..terawatt.hours.")
energy <- rename(energy, "Wind" = "Wind..terawatt.hours.")
energy <- rename(energy, "Solar" = "Solar..terawatt.hours.")
energy <- rename(energy, "Other renewables" = "Other.renewables..modern.biofuels..geothermal..wave...tidal...terawatt.hours.")
energy <- energy[ ,-2]

energy$TotalenergyUse <- rowSums(energy[, c("Hydro", "Wind", "Solar", "Other renewables")], na.rm=T)

energy <- energy[ ,-3:-6]

#Lager plot for energy world


worldenergy <- energy %>%
  filter(Country == "World")

ggplot(worldenergy, aes(x = Year, y = TotalenergyUse)) +
  geom_line()+
  ggtitle("World energy use")

#Energy Africa

africaenergy <- energy %>%
  filter(Country == "Africa")

ggplot(africaenergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("Africa energy use")

#Energy Europe

europeEnergy <- energy %>%
  filter(Country == "European Union")

ggplot(europeEnergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("European Union energy use")

#Energy Asia

asiaenergy <- energy %>%
  filter(Country == "Asia Pacific")

ggplot(asiaenergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("Asia energy use")

#Energy North America

americaenergy <- energy %>%
  filter(Country == "North America")

ggplot(americaenergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("North America Energy use")

#Energy China

chinaenergy <- energy %>%
  filter(Country == "China")

#Total energy for easy viewing

energyview <- energy



```

```{r, include=FALSE}
#Laster opp gdp percap filen og rydder den

GdpPercap <- read.csv("real-gdp-per-capita-PWT.csv")

GdpPercap <- rename(GdpPercap, "Country" = "Entity", "Year" = "Year", "GdpPercap" = "GDP.per.capita..2011.international....PPP.adjusted.US...")

GdpPercap <- GdpPercap[ , -2]

GdpUS <- GdpPercap %>%
  filter(Country == "United States")

ggplot(GdpUS, aes(x = Year, y = GdpPercap)) +
  geom_line() +
  ggtitle("North America GDP per capita")

GdpChina <- GdpPercap %>%
  filter(Country == "China")



#Tester å sette gdppercap med totalenergy use for USA. Kunne startet den lavere og på samme punkt på y aksen, men det er ingen korrelasjon uansett.
ggplot() +
  geom_line(data = GdpUS, aes(y = GdpUS$GdpPercap, x = Year)) +
  geom_line(data = americaenergy, aes(y = americaenergy$TotalenergyUse, x = Year))


```

```{r, include=FALSE}
#China alle plots

ggplot(GdpChina, aes(x = Year, y = GdpPercap)) +
  geom_line() +
  ggtitle("China GDP per capita")

ggplot(chinaenergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("China Energy use")

chinahdi <- hdi %>%
  filter(Country == "China")

ggplot(chinahdi, aes(x = Year, y = Value)) +
  geom_line() +
  ggtitle("China HDI")


```

```{r, include=FALSE}
#India, for å sjekke om populasjon kan forklare økt bruk av alternativ energi

indiaenergy <- energy %>%
  filter(Country == "India")

ggplot(indiaenergy, aes(x = Year, y = TotalenergyUse)) +
  geom_line() +
  ggtitle("India energy use")

indiaGdp <- GdpPercap %>%
  filter(Country == "India")

ggplot(indiaGdp, aes(x = Year, y = GdpPercap)) +
  geom_line() +
  ggtitle("India GdpPercap")



```


```{r, include=FALSE}
# ANIMERT GRAF AV GDP-PER-CAPITA 

gifski_renderer(file = tempfile(fileext = ".gif"), loop = TRUE,
  width = NULL, height = NULL)

file_renderer(dir = ".", prefix = "gganim_plot", overwrite = FALSE)

animplot <- ggplot(GdpPercap, aes(x=Year , y=GdpPercap, color=Country)) + 
  geom_line() + 
  geom_point() + 
  ggtitle("Gdp per capita by country from 1950 to 2018") + 
  ylab("Gdp per capita") +
  transition_reveal(Year)+
  theme_ft_rc()



```

```{r, include=FALSE}
#Uploading and fixing the population file for easy data wrangling

Population <- read.csv("population-figures-by-country-csv_csv.csv")

Population <- gather(Population, Year, Population, Year_1960:Year_2016)

Population$Year <- substring(Population$Year, 6, 9)

Population$Year <- as.numeric(Population$Year)

Population <- Population[-1:-1315, ]

Population <- Population [, -2]

#Merging it together with energyview to get Per capita

Totalpop <- merge(Population, energyview)

totalprcap <- Totalpop %>%
  group_by(Year, Country) %>%
  summarise(Prcap = TotalenergyUse/Population)

chinatest <- totalprcap %>%
  filter(Country == "China")

chinapop <- Population %>%
  filter(Country == "China")

chinatotal <- merge(chinaenergy, chinapop)

chinatotal <- chinatotal %>%
  group_by(Year) %>%
  summarise(EnergyprCap = TotalenergyUse/Population)

ggplot(chinatotal, aes(x = Year, y = EnergyprCap)) +
  geom_line()
```

```{r, include=FALSE}
#Heads (top5) for each category

Continents <- c("Non-OECD", "World", "OECD", "Asia Pacific", "Europe & Eurasia", "North America", "European Union", "South & Central America", "CIS")

#Fixing for 2016 on prcap energyuse
prcap2016 <- totalprcap %>%
  filter(Year == 2016)


Headprcap <- prcap2016[order(-prcap2016$Prcap), ]
head(unique(Headprcap$Country, 5))

#Removing continents for energyview. usinc c() didnt work so i had to force it.
energy2016 <- energyview %>%
  filter(Year == 2016)

energyview <- energyview %>%
  filter(Country != "Non-OECD") %>%
  filter(Country != "World") %>%
  filter(Country != "OECD") %>%
  filter(Country != "Asia Pacific") %>%
  filter(Country != "Europe & Eurasia") %>%
  filter(Country != "North America") %>%
  filter(Country != "European Union") %>%
  filter(Country != "South & Central America") %>%
  filter(Country != "CIS") %>%
  filter(Country != "Africa")

#Top 5 Total energy use 2016
Headtotal <- energy2016[order(-energy2016$TotalenergyUse), ] 
head(unique(Headtotal$Country, 5))

```

```{r, include=FALSE}
#Growth rate energyprcap function

growth <- function(x){(x/lag(x)-1)}

totalprcap <- totalprcap %>%
  filter(Country != "World")
totalprcap <- na.omit(totalprcap) 

rateprcap <- totalprcap %>%
  group_by(Country) %>%
  mutate_each(funs(growth), Prcap)

rateprcap <- rateprcap[-1:-54, ]

#Average growth rate, divided by 50 [(2016-1965)-1]

avgprcap <- rateprcap %>%
  group_by(Country) %>%
  summarise(AvgGrowth = sum(Prcap/50))

avgprcap <- na.omit(avgprcap)
#Average growth rate total
  
ratetotal <- energyview %>%
  group_by(Country) %>%
  mutate_each(funs(growth), TotalenergyUse)

ratetotal <- na.omit(ratetotal)

Avgtotal <- ratetotal %>%
  group_by(Country) %>%
  summarise(AvGrowth = sum(TotalenergyUse/50))

#Removing sums that divided by zero

Avgtotal <- Avgtotal %>%
  filter_all(all_vars(!is.infinite(.)))

#First 5 for rates cap and total

Headratecap <- avgprcap[order(-avgprcap$AvgGrowth), ]
head(unique(Headratecap$Country, 5))

Headtotalcap <- Avgtotal[order(-Avgtotal$AvGrowth), ]
head(unique(Headtotalcap$Country, 5))
```


```{r, include=FALSE}
#Interactive plot for total energy use
plotTotal <- energyview %>%
  SharedData$new(key = ~Country, group = "Select a country") %>%
  plot_ly(x = ~Year, y = ~TotalenergyUse, alpha = 0.5, hoverinfo = "text", text = ~Country) %>%
  group_by(Country) %>%
  add_lines() %>%
  highlight(selectize = TRUE)
plotTotal
```

```{r, include=FALSE}
#Plot for energy pr capita
plotPrcap <- totalprcap %>%
  SharedData$new(key = ~Country, group = "Select a country") %>%
  plot_ly(x = ~Year, y = ~Prcap, alpha = 0.5, hoverinfo = "text", text = ~Country) %>%
  group_by(Country) %>%
  add_lines() %>%
  highlight(selectize = TRUE)

plotPrcap
```




# Consumption of renewable energy

For our data project we wanted to look at the consumption of renewable energy. We asked ourselves what we wanted to focus on, and we decided to look closer on the total consumption in the world, and the increase in usage of renewable energy.  
  
  
* To get started, we wanted to take a look at:  
  + What are the top five countries in renewable energy consumption?
  + What are the top five countries in renewable energy consumption per capita?
  + What are the top five countries with highest average growth rate in consumption of renewable energy between 1965 and 2016?
  + What are the top five countries with highest average growth rate in consumption of renewable energy per capita between 1965 and 2016?
  + Do these countries have something in common in each of these categories?
  
  
    
      
But first we think it is wise to clarify where we got our data, and what limitations the data gave us.  

Our source is primarily Our World Data. The site is run by a subdivision on the University of Oxford, which depends on voluntary donations for its users. The site is primarily a gathering service and gathers information from other scientists, universities and independent people from all over the world, and post the data on their website. All the big data in the project has been taken from this website. The data from this site includes up to 2016, and due to some countries lacking information before 1965 we chose to start the dataset from 1965.  


After a lot of data-wrangling we manage to make some datasets that were really interesting. The first thing we looked at was the five countries that had the highest consumptions of renewable energy. From this dataset we found that china had the highest consumption in the world, almost twice as much as second place, which was USA. The rest of the list looked like this:

1. China
2. USA
3. Brazil
4. Canada
5. India

When you look at the graph below you can search for the desired country. The graph contains all the countries we had in our dataset.  
  The y-axis tells us how much energy in megawatt that where consumed in a year, and the x-axis shows each year from 1965 to 2016. We recommend looking at China, USA and Brazil to see the development differences.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
plotTotal
```

  
At first glance, we can see that the top 5 countries either has a fairly large population or a well developed economy. What we also found interesting is China being at the top, as it is also on the top of global total Co2 emissions. There are several reasons why China has expanded so much over the years in the renewable energy sector; For one, the rising population increases demand for electricity, and investing in renewable energy creates a long-lasting and safe opportunity to provide it. The second point is the health hazards currently present in China, where air pollution is the main concern, from fossil fuel and coal power plants. Hydropower is the main source of renewable energy in China, which may come as a surprise to many. With these findings, one can ask the question; Why isn't India investing nearly as much in renewable energy as China? It has the same criterias, namely a rising population, and air pollution. 

# HUSK FORKLARE RUNDT NOEN AV LANDENE EKS: USA

#### Renewable energy useage Per Capita  

  When we look at which countries has the biggest consumption per capita the list changes, only Canada is mentioned on both rankings but it has now moved to second place. Norway tops the ranking, but this is not strange at all considering that the country gets 99% of its energy from renewable energy (mainly hydropower).  
    From the table below you can see the top 5 list for energy consumption in the world, total and per capita  
      

Top 5 in the world | Top 5 per capita
------------------ | -----------------
             China | Norway
               USA | Canada
            Brazil | Sweden
            Canada | New Zealand
             India | Austria
             
From the Graph under you can choose witch country you want to look at. This graph shows you the average consumption of renewable energy per capita. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
plotPrcap
```



### Average growth-rate in the usage of renewable energy from 1965 to 2016  
  
  Now we want to look at which countries has the highest growth rate in consumption of renewable energy. The reason we wanted to look at this is because we were wondering if it is only “well developed” countries that pushes the technology forward to increase the amount of produced and used renewable energy. We made two datasets, Top 5 in the world (total) and Top 5 per capita.  
    To find which countries were in the top 5 in each category, we had to use the average growth rate. We did this to eliminate the problem where a country could have had an abnormal growth rate in 1970 and close to zero for all the other years. By doing it this way we think the data is more accurate in the representation on which country has the highest growth rate. 

  Top 5 total      | Top 5 per capita
------------------ | -----------------
      Turkmekistan | Algeria
           Algerie | Denmark
           Belarus | Hungary
           Denmark | Belgium
           Vietnam | Vietnam  
             
              
              
We looked into the sources from where the countries with high consumption of renewable energy got their energy from. All the countries got mainly their energy from hydro-alternatives, wind was second most common followed by solar. 
We then found something really surprising when we looked at the countries with high growth rate. We initially thought the countries also mainly would use hydro as their energy source, but it turned out that the these countries get their renewable energy from wind and solar. Logically, many countries may find it more beneficial to build wind turbines and solar panels. These are of course regional differences, where Denmark is a good example; Flat terrain provides lots of wind, but no slopes to build water dams for hydropower. This is where cost also comes into play: A country with no access to hydropower has to invest more into other renewable energies, which can be more costly. Example; To generate the same amount of energy, you would need alot of wind turbines rather than one water dam. This is all relative, of course, but it proves a point nonetheless. 

After some more research, we found that developing countries have poor access to formal finance witch makes it difficult to start big projects. This leads to difficulties when investors want to invest because there is a lack of investor protection. Bad social infrastructure and political instability can affect the investment poorly. Due to the limitations for investments, it appears that building structures for renewable energy may be too costly for these countries. Another interesting point is the technology difference for developing countries; Efficiency for solar panels are tied directly to new technology for batteries, etc. For the best efficiency, you need the newest technology, which proves to be costly. It would especially make sense for countries in Africa to invest in solar energy, but buying the newest technology may be too costly.

  
  
  
  
  
  
  
  
  
  

