---
title: "Final Dataproject"
author: "skyggemenn inc^TM^."
date: "24/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls())

library(dplyr)
library(ggplot2)
library(tidyverse)
library(gganimate)
library(png)
library(gifski)
library(magick)
library(broom)
library(plotly)
library(crosstalk)
packageVersion("plotly")


```

```{r, include=FALSE}
# This is the raw and unedited code we have worked with throughout the data science project. We wanted to keep it this way to show how we worked, instead of cutting it down to the bare minimum. We also included comments in case there are any codes that are not self-explanatory. There will be alot of data frames in the enviroment, but we won't contain them in lists.
```


```{r, include=FALSE}
#Laster opp og rydder HDI filen
hdi <- read.csv("human-development-index.csv")
hdi <- rename(hdi, "Country" = "Entity")
hdi <- rename(hdi, "Value" = "X..0.1..higher.values.are.better..")
hdi <- hdi[ ,-2]
```

```{r, include=FALSE}
#Laster opp og rydder renewable energy filen

energy <- read.csv("modern-renewable-energy-consumption.csv")
energy <- rename(energy, "Country" = "Entity")
energy <- rename(energy, "Hydro" = "Hydropower..terawatt.hours.")
energy <- rename(energy, "Wind" = "Wind..terawatt.hours.")
energy <- rename(energy, "Solar" = "Solar..terawatt.hours.")
energy <- rename(energy, "Other renewables" = "Other.renewables..modern.biofuels..geothermal..wave...tidal...terawatt.hours.")
energy <- energy[ ,-2]

#Summerer energiforbruket for å gjøre klar til analyse
energy$TotalenergyUse <- rowSums(energy[, c("Hydro", "Wind", "Solar", "Other renewables")], na.rm=T)

energy <- energy[ ,-3:-6]

#Lager plot for energy world


worldenergy <- energy %>%
  filter(Country == "World")

ggplot(worldenergy, aes(x = Year, y = TotalenergyUse)) +
  geom_line()+
  ggtitle("World energy use")

#Energy Africa

africaenergy <- energy %>%
  filter(Country == "Africa")

ggplot(africaenergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("Africa energy use")

#Energy Europe

europeEnergy <- energy %>%
  filter(Country == "European Union")

ggplot(europeEnergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("European Union energy use")

#Energy Asia

asiaenergy <- energy %>%
  filter(Country == "Asia Pacific")

ggplot(asiaenergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("Asia energy use")

#Energy North America

americaenergy <- energy %>%
  filter(Country == "North America")

ggplot(americaenergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("North America Energy use")

#Energy China

chinaenergy <- energy %>%
  filter(Country == "China")

#Total energy for easy viewing

energyview <- energy



```

```{r, include=FALSE}
#Laster opp gdp percap filen og rydder den

GdpPercap <- read.csv("real-gdp-per-capita-PWT.csv")

GdpPercap <- rename(GdpPercap, "Country" = "Entity", "Year" = "Year", "GdpPercap" = "GDP.per.capita..2011.international....PPP.adjusted.US...")

GdpPercap <- GdpPercap[ , -2]

GdpUS <- GdpPercap %>%
  filter(Country == "United States")

ggplot(GdpUS, aes(x = Year, y = GdpPercap)) +
  geom_line() +
  ggtitle("North America GDP per capita")

GdpChina <- GdpPercap %>%
  filter(Country == "China")



#Tester å sette gdppercap med totalenergy use for USA. Vi har prøvd mer enn bare USA, men den eneste vi kunne finne noen sammenheng med var Kina. Kunne startet den lavere og på samme punkt på y aksen, men det er ingen korrelasjon uansett.
ggplot() +
  geom_line(data = GdpUS, aes(y = GdpUS$GdpPercap, x = Year)) +
  geom_line(data = americaenergy, aes(y = americaenergy$TotalenergyUse, x = Year))


```

```{r, include=FALSE}
#China alle plots

ggplot(GdpChina, aes(x = Year, y = GdpPercap)) +
  geom_line() +
  ggtitle("China GDP per capita")

ggplot(chinaenergy, aes(x = Year, y = TotalenergyUse))+
  geom_line()+
  ggtitle("China Energy use")

chinahdi <- hdi %>%
  filter(Country == "China")

ggplot(chinahdi, aes(x = Year, y = Value)) +
  geom_line() +
  ggtitle("China HDI")


```

```{r, include=FALSE}
#India, for å sjekke om populasjon kan forklare økt bruk av alternativ energi

indiaenergy <- energy %>%
  filter(Country == "India")

ggplot(indiaenergy, aes(x = Year, y = TotalenergyUse)) +
  geom_line() +
  ggtitle("India energy use")

indiaGdp <- GdpPercap %>%
  filter(Country == "India")

ggplot(indiaGdp, aes(x = Year, y = GdpPercap)) +
  geom_line() +
  ggtitle("India GdpPercap")



```


```{r, include=FALSE}
# ANIMERT GRAF AV GDP-PER-CAPITA, kjør animplot for å se grafen.

gifski_renderer(file = tempfile(fileext = ".gif"), loop = TRUE,
  width = NULL, height = NULL)

file_renderer(dir = ".", prefix = "gganim_plot", overwrite = FALSE)

animplot <- ggplot(GdpPercap, aes(x=Year , y=GdpPercap, color=Country)) + 
  geom_line() + 
  geom_point() + 
  ggtitle("Gdp per capita by country from 1950 to 2018") + 
  ylab("Gdp per capita") +
  transition_reveal(Year)+
  theme_ft_rc()



```

```{r, include=FALSE}
#Uploading and fixing the population file for easy data wrangling

Population <- read.csv("population-figures-by-country-csv_csv.csv")

Population <- gather(Population, Year, Population, Year_1960:Year_2016)

Population$Year <- substring(Population$Year, 6, 9)

Population$Year <- as.numeric(Population$Year)

#Removing years up to 1965
Population <- Population[-1:-1315, ]

Population <- Population [, -2]

#Merging it together with energyview to get Per capita

Totalpop <- merge(Population, energyview)

totalprcap <- Totalpop %>%
  group_by(Year, Country) %>%
  summarise(Prcap = TotalenergyUse/Population)

chinatest <- totalprcap %>%
  filter(Country == "China")

chinapop <- Population %>%
  filter(Country == "China")

chinatotal <- merge(chinaenergy, chinapop)

chinatotal <- chinatotal %>%
  group_by(Year) %>%
  summarise(EnergyprCap = TotalenergyUse/Population)

ggplot(chinatotal, aes(x = Year, y = EnergyprCap)) +
  geom_line()
```

```{r, include=FALSE}
#Heads (top5) for each category

Continents <- c("Non-OECD", "World", "OECD", "Asia Pacific", "Europe & Eurasia", "North America", "European Union", "South & Central America", "CIS")

#Fixing for 2016 on prcap energyuse
prcap2016 <- totalprcap %>%
  filter(Year == 2016)


Headprcap <- prcap2016[order(-prcap2016$Prcap), ]
head(unique(Headprcap$Country, 5))

#Removing continents for energyview. usinc c() didnt work so i had to force it.
energy2016 <- energyview %>%
  filter(Year == 2016)

energyview <- energyview %>%
  filter(Country != "Non-OECD") %>%
  filter(Country != "World") %>%
  filter(Country != "OECD") %>%
  filter(Country != "Asia Pacific") %>%
  filter(Country != "Europe & Eurasia") %>%
  filter(Country != "North America") %>%
  filter(Country != "European Union") %>%
  filter(Country != "South & Central America") %>%
  filter(Country != "CIS") %>%
  filter(Country != "Africa")

#Top 5 Total energy use 2016
Headtotal <- energy2016[order(-energy2016$TotalenergyUse), ] 
head(unique(Headtotal$Country, 5))

```

```{r, include=FALSE}
#Growth rate energyprcap function

growth <- function(x){(x/lag(x)-1)}

totalprcap <- totalprcap %>%
  filter(Country != "World")
totalprcap <- na.omit(totalprcap) 

rateprcap <- totalprcap %>%
  group_by(Country) %>%
  mutate_each(funs(growth), Prcap)

rateprcap <- rateprcap[-1:-54, ]

#Average growth rate, divided by 50 [(2016-1965)-1]

avgprcap <- rateprcap %>%
  group_by(Country) %>%
  summarise(AvgGrowth = sum(Prcap/50))

avgprcap <- na.omit(avgprcap)
#Average growth rate total
  
ratetotal <- energyview %>%
  group_by(Country) %>%
  mutate_each(funs(growth), TotalenergyUse)

ratetotal <- na.omit(ratetotal)

Avgtotal <- ratetotal %>%
  group_by(Country) %>%
  summarise(AvGrowth = sum(TotalenergyUse/50))

#Removing sums that divided by zero

Avgtotal <- Avgtotal %>%
  filter_all(all_vars(!is.infinite(.)))

#First 5 for rates cap and total

Headratecap <- avgprcap[order(-avgprcap$AvgGrowth), ]
head(unique(Headratecap$Country, 5))

Headtotalcap <- Avgtotal[order(-Avgtotal$AvGrowth), ]
head(unique(Headtotalcap$Country, 5))
```


```{r, include=FALSE}
#Interactive plot for total energy use
plotTotal <- energyview %>%
  SharedData$new(key = ~Country, group = "Select a country") %>%
  plot_ly(x = ~Year, y = ~TotalenergyUse, alpha = 0.5, hoverinfo = "text", text = ~Country) %>%
  group_by(Country) %>%
  add_lines() %>%
  layout(xaxis = list(title = "Year"), yaxis = list(title = "Terrawatt hours")) %>%
  highlight(selectize = TRUE)
plotTotal
```

```{r, include=FALSE}
#Plot for energy pr capita
plotPrcap <- totalprcap %>%
  SharedData$new(key = ~Country, group = "Select a country") %>%
  plot_ly(x = ~Year, y = ~Prcap, alpha = 0.5, hoverinfo = "text", text = ~Country) %>%
  group_by(Country) %>%
  add_lines() %>%
  layout(xaxis = list(title = "Year"), yaxis = list(title = "Terrawatt hours per capita")) %>%
  highlight(selectize = TRUE)

plotPrcap
```




# What causes an increase in consumption of renewable energy?

For our data project we wanted to look at the consumption of renewable energy. We asked ourselves what we wanted to focus on, and we decided to look closer on the total consumption in the world, and the increase in usage of renewable energy.  
  
  
* To get started, we wanted to take a look at:  
  + What are the top five countries in renewable energy consumption?
  + What are the top five countries in renewable energy consumption per capita?
  + What are the top five countries with highest average growth rate in consumption of renewable energy between 1965 and 2016?
  + What are the top five countries with highest average growth rate in consumption of renewable energy per capita between 1965 and 2016?
  + Do these countries have something in common?
  
  
    
But first we think it is wise to clarify where we got our data, and what limitations the data gave us.  

Our source is primarily Our World Data. The site is run by a subdivision on the University of Oxford, which depends on voluntary donations for its users. The site is primarily a gathering service and gathers information from other scientists, universities and independent people from all over the world, and post the data on their website. All the big data in the project has been taken from this website. The data from this site includes up to 2016, and due to some countries lacking information before 1965 we chose to start the dataset from 1965. 

We tried to compare  total GDP (and per capita) against renewable energy use, but we found no correlation between the two. We also tried for HDI, but for developed countries it is mainly a linear graph. Adding more economic variables would just hinder our main objective; The problem is trying to predict a correlation rather than finding an explanation in the dataset itself.


After a lot of data-wrangling we managed to make some datasets that were quite interesting. The first thing we looked at was the five countries that had the highest consumptions of renewable energy. From this dataset we found that china had the highest consumption in the world, almost twice as much as second place, which was USA. The rest of the list looked like this:

1. China
2. USA
3. Brazil
4. Canada
5. India

When you look at the graph below you can search for the desired country. The graph contains all the countries we had in our dataset.  
  The y-axis tells us how much energy in megawatt hours that where consumed in a year, and the x-axis shows each year from 1965 to 2016. We recommend looking at China, USA and Brazil to see the development differences. China has a steady increase, until it finally gains a massive leap. United states keeps the energy use steady, but has a few dips which we will explain below. Brazil on the other hand, maintains a steady increase throughout.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
plotTotal
```

  
At first glance, we can see that the top 5 countries either has a fairly large population or a well developed economy. What we also found interesting is China being at the top, as it is also on the top of global total Co2 emissions. There are several reasons why China has expanded so much over the years in the renewable energy sector; For one, the rising population increases demand for electricity, and investing in renewable energy creates a long-lasting and safe opportunity to provide it. The second point is the health hazards currently present in China, where air pollution is the main concern, from fossil fuel and coal power plants. Hydropower is the main source of renewable energy in China, which may come as a surprise to many. With these findings, This poses an interesting question; Why isn't India investing nearly as much in renewable energy as China? It has the same criterias, namely a rising population, and air pollution! One of the answers may be that China has a "more centralized" government system that creates a better incentive to invest in specific sectors. What we found in common for these countries on top is that they have a large consumation of total energy, and a considerable percentage share of this is renewable. 
  
  ***


__Now try to search for *United States* in the graph above__. We found the US graph to be very interesting. As you can see from the graph, there are three major dips; 1973, 1980 and 2001. For 1973, this may have been caused by the 1973 energy crisis. Due to this crisis, the US decided to be energy independent and consolidated the federal energy administration which also had an energy research and development administration. All this may have led to make it more difficult to research and develop sources of renewable energy.  
In the 1980's US demand for oil started to increase, at the same time in the late 80’s the situation in the Gulf forced US to protect their oil tankers and not long after, a war broke out. We haven’t found any direct link between the build up to the war and reduction in the consumption of renewable energy, but we assume that the gulf war would have had some sort of impact.  
Then we have 2001. Since it's mainly hydro that is produced as a renewable energy in the US, it is vulnerable to drought, which happened in 2001.  

#### Renewable energy useage Per Capita  

  When we look at which countries has the biggest consumption per capita the list changes, only Canada is mentioned on both rankings but it has now moved to second place. Norway tops the ranking, but this is not strange at all considering that the country gets 99% of its energy from renewable energy (mainly hydropower).  
    From the table below you can see the top 5 list for energy consumption in the world, total and per capita  
      

Top 5 in the world | Top 5 per capita
------------------ | -----------------
             China | Norway
               USA | Canada
            Brazil | Sweden
            Canada | New Zealand
             India | Austria
             

Top 5 in the world is gathered from the dataset "energyview" while per capita is gathered from "totalprcap". (Note that the numbers in totalprcap is really small due to the original value of terrawatt hours is scaled up.)             
From the Graph below you can choose which country you want to look at. This graph shows you the consumption of renewable energy per capita. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
plotPrcap
```



### Average growth-rate in the usage of renewable energy from 1965 to 2016  
  
  Now we want to look at which countries has the highest growth rate in consumption of renewable energy. The reason we wanted to look at this is because we were wondering if it is only “well developed” countries that pushes the technology forward to increase the amount of produced and usage of renewable energy. We made two datasets, Top 5 in the world (total) and Top 5 per capita.  
    To find which countries were in the top 5 in each category, we had to use the average growth rate. We did this to eliminate the problem where a country could have had an abnormal growth rate in 1970 and close to zero for all the other years. By doing it this way we think the data is more accurate in the representation on which country has the highest growth rate. 

  Top 5 total      | Top 5 per capita
------------------ | -----------------
Turkmekistan 24.2% | Algeria 16%
    Algeria  18.6% | Denmark 15.5%
    Belarus  16.3% | Hungary 12.4%
    Denmark  15.9% | Belgium 11.5%
    Vietnam  12.6% | Vietnam 10.6%  


Top 5 total is gathered from the dataset "Headtotalcap", while Top 5 per capita is gathered from "Headratecap". 

* _Since this is the main part of our story, we want to explain in detail why these countries are on top of the list._  
  + __Algeria__  
    Algeria is the country that have highest grow rate with renewable energy. The reason for this is that they export petroleum and natural gas that is estimated to be more than 60% than their budget revenue. And over 95% of their exports earnings. Algeria went from having 80 % of their usage from oil to 60 % and 13% from Renewable energy to 26%, while the rest is from nuclear energy. Algeria has an astonishing geographic stance that gives them enormous opportunities to produce renewable energies, mainly from wind, solar and biomass. They set a goal to generate as much renewable energy as they produce from natural gas by 2020. Algeria has the potential to generate three times the necessary energy use than what they have now, All because of the geographical location of the country. The technology for generating renewable energy have expanded and made it cheaper to produce renewable energy.  
  + __Denmark__  
    43,4% of all their electricity consumption comes from wind power. It went up from 19% in 2005 to 43,4% in 2018. Denmark has a goal to have more than half their energy usage from renewable energy by 2030 and zero fossil fuel energy usage by 2050. Denmark is one of the leading countries regarding wind power generating technologies. Denmark follows a policy where electricity suppliers are obligated to produce a certain quota of renewable energy.    
  + __Hungary__  
    Hungary has good reason to improve their production of renewable energy; One of their major reasons is the political one, where Hungary has long wanted to join the EU. One aspect of EU is focusing on the renewable energy policies. For Hungary to join the EU, they have focused heavily on renewable energy. Making it one of the countries with the highest growth rate on renewable energy. Hungary was the second lowest on renewable electricity production with only 7% in EU in 2016. In 2016 Hungary sourced 18% from coal, 50% from nuclear and 20% from gas. With that low renewable energy consumption before it’s not surprising that Hungary has one of the highest growth rates with renewable energy.   
  + __Belgium__  
    One of the reasons for Belgium having a high growth rate is due to support schemes that mainly contains a quota system based on trade of certificates. Belgium also focused on wind power and hydropower with national regulations. Belgium has grown from 2% renewable energy usage to 8% from 2005 to 2014.      
  + __Vietnam__
    Vietnam has a low cost on energy consumption with 37% renewable energy usage from hydropower. The rest is mainly from fossil energy sources. Vietnam plan to increase 7% percent of renewable energy usage every year until 2020, And above 10 percent after 2020. Where reduction of energy from coal and fossil fuel is the main focus. This has made Vietnam one of the fastest growing renewable energy producers in the world.  
  + __Turkmenistan__
    Turkmenistan has an especially high prospect to produce renewable energy, due to it's geographical location. Hydropower and wind power is the two main sources where Turkmenistan can expand. With the agreement and economical support from Germany, and technology agreement with Germany, they have managed to get an enormous growth rate throughout the years.     
   + __Belarus__  
     Belarus took initiative in the UN meeting in 2007 to have a technological sharing agreement regarding among other things, renewable energy. They also have economics agreements with the developing and production of renewable energy, which has been a huge advantage. These are the main points for Belarus to become one of the most growing rate per renewable energy usage. 
     
     
### General information and findings     
              
We looked into the sources from where the countries with high consumption of renewable energy got their energy from. All the countries got mainly their energy from hydro-alternatives, wind was second most common followed by solar. 
We then found something really surprising when we looked at the countries with high growth rate. We initially thought the countries also mainly would use hydro as their energy source, but it turned out that the these countries get their renewable energy from wind and solar. Logically, many countries may find it more beneficial to build wind turbines and solar panels. These are of course regional differences, where Denmark is a good example; Flat terrain provides lots of wind, but no slopes to build water dams for hydropower. This is where cost also comes into play: A country with no access to hydropower has to invest more into other renewable energies, which can be more costly. Example; To generate the same amount of energy, you would need alot more wind turbines than one water dam. This is all relative, of course, but it proves a point nonetheless.  
  
  ***



### Regarding developing countries

After some more research, we found that developing countries have poor access to formal finance witch makes it difficult to start big projects. This leads to difficulties when investors want to invest because there is a lack of investor protection. Bad social infrastructure and political instability can affect the investment poorly. Due to the limitations for investments, it appears that building structures for renewable energy may be too costly for these countries. Another interesting point is the technology difference for developing countries; Efficiency for solar panels are tied directly to new technology for batteries, etc. For the best efficiency, you need the newest technology, which proves to be costly. It would especially make sense for countries in Africa to invest in solar energy, but buying the newest technology may be too costly.

### In summary 

Our findings indicate that an increase in renewable energy stems from several factors: 

1. Investment in renewable energy (Where a country's economy comes into play)
2. Potential for renewable energy, and regional differences (e.g slopes for dams or flat ground for wind turbines)
3. Political agreements, where a quota has to be met (e.g Paris agreement)
4. Access to technology, through neighbouring countries (Europe), or bought (China).

